\LARGE \bf Generalizing Informed Sampling for Asymptotically Optimal Sampling-based Planning via Markov Chain Monte Carlo Methods

### 2017-09-14 paper.tex
sync from sharelatex

\dy{Rohan, please finish this part}

Also we can take other approaches that generate samples from a Markov chain without requiring an approximation of a target distribution $ \pi $.
Hit-and-Run \cite{KSZ11} is one of them that efficiently generates uniform samples in convex regions.
In order to generalize our sampler to a wider range of problems~\cite{KSZ11} and better efficiency
We use the Accelerated Hit-and-Run algorithm~\cite{KSZ11}, described in Alg.~\ref{alg:hit_and_run_mcmc} that produces a series of samples $x_0, \ldots$.
Given the previous sample~$x_{i-1}$ it first  samples a random direction on the unit sphere (line 1).
This induces a line~$L(\lambda)$ passing through~$x_{i-1}$  in the direction sampled (line 2) parametrized by a scalar~$\lambda$.
We obtain upper and lower bounds on $\lambda$ (line 3) that are problem dependant. 
For example, if we have box constraints on joint limits of the robot and on maximum velocity, then bounds are given by $\lambda^{+} = -\lambda^{-} = l_{diag}$; where $l_{diag}$ is the length of the longest diagonal of the box.
We then sample a point along~$L(\lambda)$ by sampling a scalar~$\lambda'$ within our bounds (line 5).
This defines a point~$x_{i}$ which is a candidate for the next sample along the Markov Chain (line 6).
We then check if the point lies in the informed set (line 7) and if it does we return it.
If not, we update our bounds (lines~9-12) and repeat the process.
The algorithm can be viewed as an efficient method that performs rejection sampling along a one-dimensional line passing through the previous sample parametrized by $\lambda$.

@@

\os{Mention that we do not need convexity}

\dy{say some Hit-and-Run's advantages over Metropolis-Hastings}

@@

We evaluate the performance of proposed MCMC methods by comparing four types of samplers, which are rejection sampler (RS), hierarchical rejection sampler (HRS), Metropolis-Hastings sampler (MHS), and Hit-and-run sampler (HNR).
We use different samplers to generate a fixed number of samples in different informed sets to check the sampling efficiency, especially when the sampling is harder (in a small informed set).
We then evaluate how the samplers works with informed RRT*~\cite{GSB14}, which reflect the quality of samples in an informed set.

@@

\dy{TODO: change MCMC to MH (Metropolis Hasting) in the plots}

Fig. \ref{fig:sampling_efficiency} shows the log plot of the average time taken to generate one sample in the informed space vs. informed set volume ratio, i.e. the ratio of the volume of informed space to the volume of entire state space for 5000 samples. 
The informed set volume ratio is approximated by the ratio of the number of accepted to the total number samples obtained while running rejection sampling. Efficiency of rejection sampling based methods is very low for high dimesions because this ratio decreases exponentially w.r.t the dimension of the space due to the curse of dimensionality. 
The Fig. \ref{fig:sampling_efficiency} shows that MCMC and Hit\&Run have a better sampling efficiency compared to HRS and RS with decrease in informed set ratio or increase in dimensions. 
Figure \ref{fig:sampling_efficiency:levelset} shows how the informed set volume ratio decreases as the level set cost $ c_{best} $ becomes smaller in problems of different dimensions.
In a higher dimension, the informed set volume ratio falls much more quickly as the level set cost $ c_{best} $ decreases in planning process, as new better solutions are kept being found.

@@

Metropolis-Hastings shows consistent sampling time when problems get harder. 
It takes the advantage of sampling a near state that generate samples in a level set.
However this does not reflect the quality of the samples, though all the samples are in the level set.
Recall in Algorithm~\ref{alg:mh_mcmc}, a new sample candidate is obtained from a Gaussian distribution $ q( x \mid x_{i-1}, \Sigma ) $.
The best covariance $ \Sigma $ that generates faster convergence differs with problem setting.
A small covariance tends to generate more samples near around previous samples, which a large covariance is more likely to drive a Markov chain outside of an informed set.
In our next planning experiment setting, we use the same covariance for different problems.

@@

We want to point out that efficiently sampling in an informed set is not sufficient for determining the quality of a sampler.
For example, a sampler that constantly only return the same sample in a level set might show the best sampling efficiency can be the worst sampler in a path planning problem.
\dy{is the property that any state in a set has a chance to be sampled?}
We hope sampler (satisfies some properties ?) that a resulting planner is asymptotically optimal.

@@

To evaluate how the samplers impact the planning efficiency, we run the samplers with the informed RRT* planner~\cite{GSB14} on three different problems described below and shown in Fig. \ref{fig:problems}. 
For each problem the start and goal states (positions and velocities) are known in the joint space.

\subsubsection{Problem 1: 3 DoF Planar Manipulator}

Problem 1 is a 3 DoF manipulator that moves on a plane, which results in a 6 dimension space.
The position boundary of each join is $ [ - \pi , \pi ] $, the velocity boundary is $ [-10, 10] $, and the acceleration is $ 1 $.
Figure~\ref{fig:planning_efficiency:3dof:example} shows a few steps of a planned path.

@@

Problem 2 is a 6 DoF snake arm that moves in a 3D workspace, which results in a 12 dimension space.
The position boundary of each join is $ [ - \pi , \pi ] $, the velocity boundary is $ [-10, 10] $, and the acceleration is $ 1 $.
Figure~\ref{fig:planning_efficiency:6dof:example} shows a few steps of a planned path.

@@

Problem 3 is a 7 DoF Barrett WAM arm of HERB that moves in a 3D workspace, which results in a 14 dimension space.
The position boundaries of 7 joints of the arm are $  [0.541593 , 5.74159] $, $  [-2 , 2] $, $ [-2.8 , 2.8] $, $ [-0.9 , 3.1] $, $ [-4.76 , 1.24] $, $ [-1.6 , 1.6] $ and $ [-3 , 3] $.
The velocity boundary of 7 joints of the arm are $ [-0.75 , 0.75] $, $ [-0.75 , 0.75] $, $ [-2 , 2] $, $ [-2 , 2] $, $ [-2.5 , 2.5] $, $ [-2.5 , 2.5] $ and $ [-2.5 , 2.5] $.
The accelerations of all the joints are set to $ 1 $.
Figure~\ref{fig:planning_efficiency:herb:example} shows a few steps of a planned path.

@@

HNR shows close performance with RS and HRS in a 6-dimension space, as in Figure~\ref{fig:planning_efficiency:3dof:general}.
But in higher dimension performance, the advantage of HNR in sampling efficiency is shown.
In Figure~\ref{fig:planning_efficiency:6dof:hammering} and \ref{fig:planning_efficiency:herb:batting}, the cost of current best solution by planners with HNR sampler converge significantly faster.
It shows that HNR will have close performance with RS and HRS in simple problem, as in Figure \ref{fig:planning_efficiency:3dof:general}.
But when the dimensions of problems increase, HNR shows better and steady performance.
Also, HNR also show fast convergence to the optimal, because it is better at adapting the informed set update in planning process.

>>>>

Hit-and-Run~\cite{S84} MCMC sampler is known to efficiently generate uniform samples. Specifically, we use the Accelerated Hit-and-Run variant~\cite{KSZ11} of the algorithm i.e. described in Alg.~\ref{alg:hit_and_run_mcmc}.
Given the previous sample~$x_{i-1}$ it first  samples a random direction on the unit sphere (line 1).
This induces a line~$L(\lambda)$ passing through~$x_{i-1}$ in the direction sampled (line 2), parametrized by a scalar~$\lambda$.
We obtain upper and lower bounds on $\lambda$ (line 3) that are problem dependant. 
For example, if we have box constraints on joint limits of the robot and on maximum velocity, then bounds are given by $\lambda^{+} = -\lambda^{-} = l_{diag}$; where $l_{diag}$ is the length of the longest diagonal of the box.
We then sample a point along~$L(\lambda)$ by sampling a scalar~$\lambda'$ within our bounds (line 5).
This defines a point~$x_{i}$ which is a candidate for the next sample along the Markov Chain (line 6).
We then check if the point lies in the informed set (line 7) and if it does, we return it.
If not, we update our bounds (lines~9-12) and repeat the process.
The algorithm can be viewed as an efficient method that performs rejection sampling along a one-dimensional line passing through the previous sample parametrized by $\lambda$.

@@

The Hit-and-Run algorithm is designed specifically for uniform distributions. 
Hence, it doesn't require an approximation of a target distribution $ \pi $ that was generated with the energy function for Metropolis-Hastings algorithm.

@@

We evaluate the performance of proposed MCMC methods by comparing four types of samplers, which are Rejection Sampler (RS), Hierarchical Rejection Sampler (HRS), Metropolis-Hastings Sampler (MHS), and Hit-and-Run Sampler (HNR).
We use different samplers to generate a fixed number of samples in different informed sets to check the sampling efficiency. %, especially when the sampling is harder (in a small informed set).
We then compare the quality of the samplers by evaluating how the samplers work with informed RRT*~\cite{GSB14}.

@@

\dy{TODO: change MCMC to MHS (Metropolis Hasting Sampler) in the plots}

Fig. \ref{fig:sampling_efficiency} shows the log plot of the average time taken to generate one sample in the informed space vs. informed set volume ratio, i.e. the ratio of the volume of informed space to the volume of entire state space, for 5000 samples. 
The informed set volume ratio is approximated by the ratio of the number of accepted to the total number of samples obtained while running rejection sampling. 
%Efficiency of rejection sampling based methods is very low for high dimensions because this ratio decreases exponentially w.r.t the dimension of the space, for a given sub-level-set of the cost function. 
Fig. \ref{fig:sampling_efficiency} shows that MHS and HNR have a better sampling efficiency compared to HRS and RS with decrease in informed set volume ratio or increase in dimensions. 
Figure \ref{fig:sampling_efficiency:levelset} shows how the informed set volume ratio decreases as the level set cost $ c_{best} $ becomes smaller in problems of different dimensions.
In a higher dimension, the informed set volume ratio falls much more quickly as the level set cost $ c_{best} $ decreases in planning process, as new better solutions are kept being found.

@@

Metropolis-Hastings shows consistent sampling time when problems get harder. 
It takes the advantage of sampling a near state that generate samples in a level set.
However this does not reflect the quality of the samples, though all the samples are in the level set.
Recall in Algorithm~\ref{alg:mh_mcmc}, a new sample candidate is obtained from a Gaussian distribution $ q( x \mid x_{i-1}, \Sigma ) $.
The best covariance $ \Sigma $ that generates faster convergence differs with problem setting.
A small covariance tends to generate more samples near around previous samples, which a large covariance has better exploration but is more likely to drive a Markov chain outside of an informed set.
In our next planning experiment setting, we use the same covariance for different problems.

@@

We want to point out that efficiently sampling in an informed set is not sufficient for determining the quality of a sampler.
For example, a sampler that constantly returns the same sample in a informed set might show the best sampling efficiency, however it is the worst sampler in a path planning problem.
\dy{is the property that any state in a set has a chance to be sampled?}
We hope sampler (satisfies some properties ?) that a resulting planner is asymptotically optimal.

@@

To evaluate how the samplers impact the planning efficiency, we run the samplers with the informed RRT* planner~\cite{GSB14} on three different problems described below and shown in Fig. \ref{fig:problems}. 
For each problem the start and goal states (positions and velocities) are known in the joint space.
\dy{TODO: Make table for limits and add units}
\subsubsection{Problem 1: 3 DoF Planar Manipulator}

Problem 1 is a 3 DoF manipulator that moves on a plane, which results in a 6 dimension space.
The position boundary of each join is $ [ - \pi , \pi ] $, the velocity boundary is $ [-10, 10] $, and the acceleration is $ [-1,1] $.
Figure~\ref{fig:planning_efficiency:3dof:example} shows a few steps of a planned path.

@@

Problem 2 is a 6 DoF snake arm that moves in a 3D workspace, which results in a 12 dimension space.
The position boundary of each join is $ [ - \pi , \pi ] $, the velocity boundary is $ [-10, 10] $, and the acceleration is $ [-1,1] $.
Figure~\ref{fig:planning_efficiency:6dof:example} shows a few steps of a planned path.

@@

Problem 3 is a 7 DoF Barrett WAM arm of HERB that moves in a 3D workspace, which results in a 14 dimension space.
The position boundaries of 7 joints of the arm are $  [0.541593 , 5.74159] $, $  [-2 , 2] $, $ [-2.8 , 2.8] $, $ [-0.9 , 3.1] $, $ [-4.76 , 1.24] $, $ [-1.6 , 1.6] $ and $ [-3 , 3] $.
The velocity boundary of 7 joints of the arm are $ [-0.75 , 0.75] $, $ [-0.75 , 0.75] $, $ [-2 , 2] $, $ [-2 , 2] $, $ [-2.5 , 2.5] $, $ [-2.5 , 2.5] $ and $ [-2.5 , 2.5] $.
The accelerations of all the joints are set to $ [-1,1] $.
Figure~\ref{fig:planning_efficiency:herb:example} shows a few steps of a planned path.

@@

HNR shows close performance with RS and HRS in a 6-dimension space, as in Figure~\ref{fig:planning_efficiency:3dof:general}.
But in higher dimension performance, the advantage of HNR in planning efficiency is shown.
In Figure~\ref{fig:planning_efficiency:6dof:hammering} and \ref{fig:planning_efficiency:herb:batting}, the cost of current best solution by planners with HNR sampler converge significantly faster.
It shows that HNR will have close performance with RS and HRS in simple problem, as in Figure \ref{fig:planning_efficiency:3dof:general}.
But when the dimensions of problems increase, HNR shows better and steady performance.
HNR also shows faster convergence to the optimal trajectory, because it is better at adapting the informed set update in planning process.
